The difference operation between templates of binary cellular automata

### 2015-12-27 manuscript/article.tex
Camera-ready format

\documentclass{llncs}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx} \graphicspath{ {../img/} }

\begin{document}
\pagestyle{headings}  % switches on printing of running heads
\addtocmark{The difference operation between templates of binary cellular automata} % additional mark in the TOC

@@

% \author{Zorandir Soares\inst{2} \and Maurício Verardo\inst{2} \and
% Pedro P. B. de Oliveira\inst{1}\inst{2}}

% \authorrunning{Zorandir Soares, Maurício Verardo, and Pedro de Oliveira}

% \tocauthor{Zorandir Soares, Maurício Verardo, and Pedro de Oliveira}

% \institute{Universidade Presbiteriana Mackenzie\\
% Faculdade de Computação e Informática
% \and
% Pós-Graduação em Engenharia Elétrica e Computação\\
% Rua da Consolação 896, Consolação\\
% 01302-907 São Paulo, SP - Brazil\\
% \email{zorandir@gmail.com}\\
% \email{mauricio.verardo@gmail.com}\\
% \email{pedrob@mackenzie.br}}

@@

\begin{abstract}
A cellular automata template is a generalisation of the state transition tables of cellular automata. Templates have recently been introduced to represent families of cellular automata that share a given property. This paper introduces the operation of difference between templates, used to find cellular automata that have a given property but lack another. We also introduce a process called \textit{exception templates} that is required to this operation. Experimental results of both techniques are illustrated with examples in the space of elementary cellular automata.
\keywords{Cellular automata, rule space, templates, difference between templates, exception templates.}
\end{abstract}
%
\section{Introduction}
\label{sec:introducao}
Cellular automata (CAs) are dynamical systems typically discrete in time, space and state variables.
CAs can produce behavior of great complexity even based on simple rules of local action \cite{wolfram2002}. The study of classical problems of cellular automata, like the parity problem \cite{Betel2013} and the density classification problem \cite{deOliveira2014density} can help understand how this complex behavior emerges.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{fig_ruleIcon60.png}
  \caption{Rule table of the elementary CA 60.}
  \label{fig:table60}
\end{figure}

@@

%\subsection{Internal symmetry}
Another interesting property is internal symmetry. But in order to grasp its meaning, first we need the concept of rule equivalence. The following explanations are valid for binary rules, although it is possible to generalise the notion to arbitrary $k$ states.
  \begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{fig_ruleIcon102.png}
    \caption{Elementary CA 102, equivalent by reflection to elementary CA 60.}
    \label{fig:table102}
  \end{figure}

@@

% ---- Bibliography ----
\bibliography{bibliografia}
\bibliographystyle{splncs}

\end{document}

>>>>

\documentclass{llncs}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}

\begin{document}
\pagestyle{headings}
\addtocmark{The difference operation between templates of binary cellular automata}

@@

\author{Zorandir Soares\inst{2} \and Maurício Verardo\inst{2} \and
Pedro P. B. de Oliveira\inst{1}\inst{2}}

\authorrunning{Zorandir Soares, Maurício Verardo, and Pedro de Oliveira}

\tocauthor{Zorandir Soares, Maurício Verardo, and Pedro de Oliveira}

\institute{Universidade Presbiteriana Mackenzie\\
Faculdade de Computação e Informática
\and
Pós-Graduação em Engenharia Elétrica e Computação\\
Rua da Consolação 896, Consolação\\
01302-907 São Paulo, SP - Brazil\\
\email{zorandir@gmail.com}\\
\email{mauricio.verardo@gmail.com}\\
\email{pedrob@mackenzie.br}}

@@

\begin{abstract}
A cellular automata template is a generalisation of the state transition tables of cellular automata. Templates have recently been introduced to represent families of cellular automata that share a given property. This paper introduces the operation of difference between templates, used to find cellular automata that have a given property but lack another. We also introduce a process called \textit{exception templates} that is required to this operation. Experimental results of both techniques are illustrated with examples in the space of elementary cellular automata.
\keywords{Cellular automata, rule space, templates, difference between templates, exception templates.}
\end{abstract}

\section{Introduction}
\label{sec:introducao}
Cellular automata (CAs) are dynamical systems typically discrete in time, space and state variables.
CAs can produce behavior of great complexity even based on simple rules of local action \cite{wolfram2002}. The study of classical problems of cellular automata, like the parity problem \cite{Betel2013} and the density classification problem \cite{deOliveira2014density} can help understand how this complex behavior emerges.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{Figure1.pdf}
  \caption{Rule table of the elementary CA 60.}
  \label{fig:table60}
\end{figure}

@@

Another interesting property is internal symmetry. But in order to grasp its meaning, first we need the concept of rule equivalence. The following explanations are valid for binary rules, although it is possible to generalise the notion to arbitrary $k$ states.
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{Figure2.pdf}
    \caption{Elementary CA 102, equivalent by reflection to elementary CA 60.}
    \label{fig:table102}
  \end{figure}

@@

% ---- Bibliography ----
\begin{thebibliography}{10}
\bibitem{wolfram2002}
Wolfram, S.:
\newblock A new kind of science. Volume~5.
\newblock Wolfram media Champaign (2002)

\bibitem{Betel2013}
Betel, H., {De Oliveira}, P.P.B., Flocchini, P.:
\newblock Solving the parity problem in one-dimensional cellular automata.
\newblock Natural Computing \textbf{12} (2013)  323--337

\bibitem{deOliveira2014density}
{De Oliveira}, P.P.B.:
\newblock On density determination with cellular automata: Results,
  constructions and directions.
\newblock Journal of Cellular Automata \textbf{9} (2014)  357--385

\bibitem{wolz2008very}
Wolz, D., {De Oliveira}, P.P.B.:
\newblock Very effective evolutionary techniques for searching cellular
  automata rule spaces.
\newblock J. Cellular Automata \textbf{3} (2008)  289--312

\bibitem{deOliveira2014}
{De Oliveira}, P.P.B., Verardo, M.:
\newblock Representing families of cellular automata rules.
\newblock The Mathematica Journal \textbf{16} (2014)

\bibitem{deOliveira2014b}
{De Oliveira}, P.P.B., Verardo, M.:
\newblock Template based representation of cellular automata rules.
\newblock In Isokawa, T., Imai, K., Matsiu, N., Peper, F., Umeo, H., eds.: 20th
  International Workshop on Cellular Automata and Discrete Complex Systems,
  Himeji, Japão, Julho 7-9 (2014)  199--204

\bibitem{wolfram1994cellular}
Wolfram, S.:
\newblock Cellular automata and complexity: collected papers. Volume~1.
\newblock Addison-Wesley Reading (1994)

\bibitem{boccara2002}
Boccara, N., {Fuk\'{s}}, H.:
\newblock Number-conserving cellular automaton rules.
\newblock Fundamenta Informaticae \textbf{52} (2002)  1--13

\bibitem{salo2013color}
Salo, V., T{\"o}rm{\"a}, I.:
\newblock Color blind cellular automata.
\newblock In: Cellular Automata and Discrete Complex Systems.
\newblock Springer (2013)  139--154

\bibitem{CATemplates}
Verardo, M., {De Oliveira}, P.P.B.:
\newblock CATemplates. (2015) https://github.com/mverardo/CATemplates.

\end{thebibliography}
\end{document}
